A Coq Formalization of Stable Sorts, Radix Sort, and the Burrows-Wheeler Transform

### 2019-10-20 draft_paper/draft_paper.tex
Rephrase intro

There has not been much work on formally verifying stable sorts, or
rather, the stability of stable sorts. Some work has been done on
verifying the stability of mergesort
\cite{leroy_2018,Sternagel2013,Leino2015,deGouw2014}, but we do not
know of any formally verified implementation of a radix sort that
includes stability.

In addition, to our knowledge no one has produced a formally verified
implementation of the BWT. Bird et~al.\ \cite{alg-of-prog,birdmu}
provide an excellent derivation of the BWT and its inverse based on
the technique of program calculation, in which programs are calculated
from their specifications using algebraic transformations. However,
their derivation leaves a few crucial lemmas unproven, including the
correctness of LSD radix sort.

In this paper, we develop more fully the theory of stable sorts: we
give multiple definitions for permutations and stable permutations and
prove them equivalent. This allows us to choose whichever definition
makes the proof easiest. We also prove that any two sorts are
interchangeable.
\begin{restatable*}[StableSort\_unique]{corollary}{stablesortunique}
  \label{thm:stablesort_unique}
  For all stable sorting functions $f, g: \var{list}~A \to
  \var{list}~A$,
  \begin{equation*}
    (\forall l,f~l = g~l)
  \end{equation*}
\end{restatable*}

>>>>

Bird et~al.\ \cite{alg-of-prog,birdmu} provide an excellent derivation
of the BWT and its inverse based on the technique of program
calculation, in which programs are calculated from their
specifications using algebraic transformations. However, their
derivation leaves a few crucial lemmas unproven, including the
correctness of LSD radix sort.

Although there is a line of work on mergesort stability proofs
\cite{leroy_2018,Sternagel2013,Leino2015,deGouw2014}, these proofs are
not directly applicable to radix sort. In fact, the correctness of LSD
radix sort remains unproven and more generally, nobody has contributed
a general theory for reasoning about stable sorts.

In this paper, we develop such a theory: we give multiple definitions
for permutations and stable permutations and prove them equivalent.
This allows us to choose whichever definition makes the proof easiest.
We also prove that any two sorts are interchangeable.
\begin{restatable*}[StableSort\_unique]{corollary}{stablesortunique}
  \label{thm:stablesort_unique}
  For all stable sorting functions $f, g: \var{list}~A \to
  \var{list}~A$,
  \begin{equation*}
    (\forall l,f~l = g~l)
  \end{equation*}
\end{restatable*}
